---
title: "Code of the statistical analysis for the article 'Life Beyond A Jar: Effects of Tank Size and Enrichment on the Behaviour and Welfare of Siamese Fighting Fishes'"
date: "`r Sys.Date()`"
author: "Juliette Tariel-Adam"
format: 
   html:
    embed-resources: true
    self-contained-math: true
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk:
    fig-align: center
    out-width: "100%"
---

```{r}
#| include: false
# Run the scripts that load the data and that contains the functions to plot and do the linear models
source("Script/Load-data.R")
source("Script/Functions/Plots.R")
source("Script/Functions/Linear-model.R")
```

The script contains the code of data transformation, analysis and graphs for the article "Life Beyond A Jar: Effects of Tank Size and Enrichment on the Behaviour and Welfare of Siamese Fighting Fishes" by XXauthorsXXX. It is in two versions: the raw R script in qmd format and the code output in html format. 

The goal was to test the effect of tank enrichment and size on Siamese fish behaviour. The tanks Jar, Small, Medium and Large are compared among each other, and the tank Large and Barren between each other. Barren is not compared to Jar, Small or Medium because they differ by two factors: enrichment and size. 

On the html version of the code, you fill find below a first row of tabs to click on (PCA, Swimming, etc) and for most of these tabs, a second row of tabs to click on (Distribution, Plot, etc).

::: {.panel-tabset}

# Data description

After staying 3-7 days in a tank, the behaviour of all fish were scored by four 10-min trials in their tank. All trials occurred on the same day at 4 different times (7 am, 10 am, 2 pm, 6 pm). During the 10 min trial = 600 sec, the fish's behaviour was assigned to one of the 10 behaviours, meaning that the behaviours are mutually exclusive (for instance, a fish can't be scored as nest building and interacting with the surface at the same time): Resting, Swimming, Hovering, Sinking/Floating, Stereotypic swimming, Nest building, Foraging, Interation with surface, Out of view, Unsure. Unsure is not present in the raw dataset.
Out.of.view is a bit special because it is not a behaviour and was not analysed.
Sinking/Floating was not analysed (see explanation in the PCA section).

::: {.panel-tabset}

## Basic description

The data has been inspected and corrected for data entry errors (sum of scored behaviours too low/high, column shifted, typos, inconsistent nb of observations per fish/time/tank...).

Each row represents a trial. For each trial, there is:
+ Fish: the fish tested
+ Tank: in which tank the fish was
+ Order: the tank order, i.e. whether it was the first, second, ..., fifth tank that the fish experienced.
+ Time: time of the day at which the trial happened
+ Filter: where there a filter in the tank
+ Behavioural columns. The columns for the behaviours are the time spent doing the specific behaviour in seconds during the trial. Some behaviours (Forgaging, Hovering, Interaction.with.Surface, Stereotypic.swimming) have been trasnformed in binary variable (.bin). The columns .bins represents if the fish performed the specific behaviour at all during the trial or not. 

```{r}
#| warning: false
summary(data)

# Number of observations per fish per tank
data %>% group_by(Fish, Tank) %>%
  summarise(n=n(), .groups = 'drop') %>%
  spread(Tank,n) 

# Number of observations per fish per time
data %>% group_by(Fish, Time) %>%
  summarise(n=n(), .groups = 'drop') %>%
  spread(Time,n) 

# Sum of all behaviours per trial
data3$total <- rowSums(dplyr::select(data3, beh.cols3)) 
sort(data3$total) 
```

The sum of all behaviours is not exactly 600 sec (as Unsure is not in the dataset).


## Plots

The mean value over multiple trials is plotted to only have one value per tank and per behaviour. 

```{r}
#| column: page
#| echo: false
#| warning: false
# the plot1 function is in the script "Plots.R"
plot1("Tank", main = data3, beh_cols = beh.cols3, palette = palette.beh3)

plot1("Tank","Fish", main = data3, beh_cols = beh.cols3, palette = palette.beh3) +  facet_wrap_paginate(~Fish,ncol=2,nrow=2, page = 1) + ggtitle("Per fish") 
plot1("Tank","Fish", main = data3, beh_cols = beh.cols3, palette = palette.beh3) +  facet_wrap_paginate(~Fish,ncol=2,nrow=2, page = 2) + ggtitle("Per fish")
plot1("Tank","Fish", main = data3, beh_cols = beh.cols3, palette = palette.beh3) +  facet_wrap_paginate(~Fish,ncol=2,nrow=2, page = 3) + ggtitle("Per fish")
plot1("Tank","Fish", main = data3, beh_cols = beh.cols3, palette = palette.beh3) +  facet_wrap_paginate(~Fish,ncol=2,nrow=2, page = 4) + ggtitle("Per fish")

plot1("Time", main = data3, beh_cols = beh.cols3, palette = palette.beh3) + ggtitle("Per time")

plot1("Order", main = data3, beh_cols = beh.cols3, palette = palette.beh3) + ggtitle("Tank order")
```

## Filter

```{r}
# nb of fish had a filter in their tank
table(data$Tank, data$Filter)/4 
# nb of times each fish had a filter in their tank
table(data$Fish, data$Filter)/4 

plot1("Filter", main = data3, beh_cols = beh.cols3, palette = palette.beh3)
```

There have never been a filter in Jar or Small. Some fish never had a filter in their tank. Filter needs to be included in all models as a control variable.

::: 

# PCA

We run a PCA in order to:

+ Try to reduce the number of variables to analyse 
+ Determine which behaviours were important to explain the variability between trials
+ See correlations between behaviours

The PCA did not help reduce the number of variables to analyse (see Analysis_PCA.html for more explanation) so the behaviours were analysed separately with linear models. The advantage of separate linear models compared to the PCA is the possibility to quantitatively interpret the differences between tanks.

Even if we did not use PCA for further analysis, the first 4 components are plotted below, which explained 71% of variance in the data, to show the important variables and the correlations between variables. 

```{r}
plot(pca, choix = "var") + labs(title = NULL)
plot(pca, choix = "var", axes = c(3,4)) + labs(title = NULL)
```

```{r}
# Contribution to each principal component
round(pca$var$cos2,2)
# Extent to which each behavioural type contributes to the total variance explained by the first 3 components
rowSums(pca$var$contrib[,1:3])/3
```

+ The 1st PCA component was driven by Resting (cos2 = 0.8), Swimming (0.7) and Foraging (0.3). Swimming and Foraging opposite to Resting. This first component could be interpret as "activity". If a fish spent a lot of time swimming and foraging during a trial, it was very little resting. This first component explained 25% of the variance in the data (as a reminder, the data is the time spent performing the different types of behaviour). So 1 quarter of the variance is a matter of activity. Keep in mind for the rest of the analyses that these behavioural types are correlated. 

+ The 2nd PCA component was driven by Hovering (cos2 = 0.5) and Stereotypic.swimming (0.5), and a bit by Interaction with surface (0.16). Hovering and Interacting.with.Surface opposite to Stereotypic.swimming. If the fish spent a lot of time hovering (and interacting with the surface), it was very little stereotypic swimming.

+ The 3rd PCA component was driven mainly by Nest building (cos2 = 0.6), and a bit by Foraging (0.16), Resting (0.14) and Stereotypic Swimming (0.13). Foraging and Resting opposite to Nest building and Stereotypic Swimming. If a fish spent time foraging and resting in a trial, it was little nest building and stereotypic swimming. This makes sense, it a fish is building a nest,it is an important task and needs to be finished, so it is not scattering its time with other activities. The apparent correlation between nest building and stereotypic swimming could be due to Jar where fish in this tank rarely engaged in nest building or stereotypic swimming. 

+ The 4th PCA component was not easily interpretable = not easily linked to the initial behavioural variables.

Sinking/Floating contributed almost nothing to the PCA (1.8%) + did not correlate strongly with any of the component. Sinking/Floating was not important to explain the variability between trials. We did not further analyse this behaviour.

# Swimming

::: {.panel-tabset}

## Distribution 

```{r}
hist1(data$Swimming)
```

## Plot 

The first plot is to see the trend by tank. Big red dots is for the overall mean  by tank. Small black points are raw data = swimming time of each trial. 

The second plot is to see the trend by fish and by tank. The dashed black line is means by tank. The other lines are the means for each fish by tank. 

```{r}
#| warning: false
plot_var(data, "Swimming")
plot_var_fish(data, "Swimming")
```

## Linear model 

```{r}
#| column: page
mSwimming <- lmer(Swimming ~ Tank + Time + Filter + Order + (1|Fish), data = data)
# Error message due to the fact that all trials with Tank = Barren are also Order = "5"
anova(mSwimming, ddf = "Kenward-Roger", type = 2)
contr_tank.size(mSwimming)
contr_tank.enrich(mSwimming)
# estimated means of time spent swimming depending on:
emmeans(mSwimming, ~ Time) # time of day
emmeans(mSwimming, ~ Filter) # filter
rbind(emmeans(mSwimming, ~ Order, at = list(Tank = c("Jar", "Small", "Medium", "Large"),Order = c("1", "2", "3", "4"))),
      emmeans(mSwimming, ~ Order, at = list(Tank = c("Barren"),Order = "5"))) # tank order
```

Effect of Tank, Time, Filter and Order! 

+ Tank. Large swam significantly more time than Jar, Small and Medium. Fish in Large swan 92 sec [CI 46, 138] more than Jar (over a 600 sec trial), xx sec [CI xx, xx] more than Small, and xx sec [CI xx, xx] more than Medium. In addition,  fish in Large swam significantly more time (xx sec [CI xx, xx]) than fish in Barren. 
+ Time. Fish swim more in the morning than in the afternoon, especially at 7 am.
+ Filter. Fish swim more with a filter in their tank.
+ Order. All confidence intervals are overlapping so probably no significant difference between orders. It seems that fish swim more in their 2nd and 3rd tank.

+ Effect of Barren or tank order 5? The effect size of Tank (shown by the Sum Sq) is three times more than the effect size of Order. Max difference between orders 1-4 is 43 sec. Difference between Large and Barren/5 is 53. Effect of tank is stronger on time spent swimming, so it is likely the diff between Large and Barren/5 is due (at least party) to Tank.

## Between-fish diff

```{r}
# Uncomment the 2 lines below if running the script for the first time
## rpt.swimming <- rpt(Swimming ~ Tank + Time + Filter + Order + (1|Fish), grname = "Fish", data = data, datatype = "Gaussian", nboot = 1000, npermut = 1000)
## save(rpt.swimming, file = "Script/Repeatability rptR output/rpt.swimming")
base::load("Script/Repeatability rptR output/rpt.swimming")
summary(rpt.swimming)
```

Likelihood ratio test indicates a statistically significant random effect. Meaning that fish behaved differently among each others in terms of swimming. But the repeatability is quite low (0.2 [95% CI 0.06, 0.37]). 

## Diagnosis

```{r}
lm_diagnosis(mSwimming)
```

The model seems good.

:::

# Resting

::: {.panel-tabset}

## Distribution 

```{r}
hist1(data$Resting)
table(data$Resting==0)
```

Only 11 trials where the fish did not rest at all. 

## Plot 

```{r}
#| warning: false
plot_var(data, "Resting")
plot_var_fish(data, "Resting")
```

## Linear model 

```{r}
#| column: page
mResting <- lmer(Resting ~ Tank + Time + Filter + Order + (1|Fish), data = data)
anova(mResting, ddf = "Kenward-Roger", type = 2)
contr_tank.size(mResting)
contr_tank.enrich(mResting)
emmeans(mResting, ~ Time)
rbind(emmeans(mResting, ~ Order, at = list(Tank = c("Jar", "Small", "Medium", "Large"),Order = c("1", "2", "3", "4"))),
      emmeans(mResting, ~ Order, at = list(Tank = c("Barren"),Order = "5")))
```

Effect of Tank, Time and Order.

+ Tank. Large rested significantly less time than Jar and Medium (almost significant for Small). Fish in Large rested 109 sec [CI 40, 178] **less** time than fish in Jar (over a 600 sec trial), and xx sec [CI xx, xx] more  fish in Medium. This is in line with the PCA and the results of Swimming: Fish that spent time swimming spent less time resting. Fish in Large were overall more active. No significant difference between Large and Barren. 
+ Time. Fish rest less time in the morning, especially at 7 am. In line with the PCA and the results of Swimming.
+ Order. All confidence intervals are overlapping so probably no significant difference between orders. It seems that fish rest more in their 1st and 4th tank.

## Between-fish diff

```{r}
# Uncomment the 2 lines below if running the script for the first time
## rpt.resting <- rpt(Resting ~ Tank + Time + Filter + Order + (1|Fish), grname = "Fish", data = data, datatype = "Gaussian", nboot = 1000, npermut = 1000)
## save(rpt.resting, file = "Script/Repeatability rptR output/rpt.resting")
base::load("Script/Repeatability rptR output/rpt.resting")
summary(rpt.resting)
```

Likelihood ratio test indicates a statistically significant random effect. Meaning that fish behaved differently among each others in terms of swimming. The repeatability is good 0.4 [95% CI 0.16, 0.58].

## Diagnosis

```{r}
lm_diagnosis(mResting)
```

The model seems good.

## Resting place

Just a plot with the amount of time resting at the different places depending on the tank. Each bar value is the mean of resting times over trials and fish. 

Analysis only for Small, Medium and Large.

```{r}
#| column: page
#| echo: false
#| message: false
data_RP %>% group_by(Tank, Resting.place) %>% summarise(Resting = mean(Resting)) %>% 
  ggplot(., aes(y = Resting, x = Tank,fill = Resting.place))+
    geom_bar(position = position_fill(reverse = TRUE), stat = "identity", color = "black")+
  scale_fill_manual(values=palette_RP)+
  guides(fill = guide_legend(reverse = TRUE))+
  ylab("% of time resting at the different places")+
  scale_y_continuous(labels=c(0,25,50,75,100))

data_RP %>% group_by(Fish, Resting.place) %>% summarise(Resting = mean(Resting)) %>% 
  ggplot(., aes(y = Resting, x = Fish,fill = Resting.place))+
    geom_bar(position = position_fill(reverse = TRUE), stat = "identity", color = "black")+
  scale_fill_manual(values=palette_RP)+
  guides(fill = guide_legend(reverse = TRUE))+
  ylab("% of time resting at the different places")+
  scale_y_continuous(labels=c(0,25,50,75,100))
```

From the graphs, we can see that fish use different places to rest, not only one type. They use all the different places available. It does not seem that there is a preferred resting place. 

Interesting, fish avoided resting on filter. 48 trials with a filter in the tank (24 in Medium + 24 in Large) but only one trial were the fish rested on the filter for Medium and Large. Fish only rested on filter in Barren. 

```{r}
# Number of trials with a filter
table(data$Tank, data$Filter)
# Trials with resting on filter
filter(master[,c(1:5,7:8)], Resting.place == "Filter")  
```


:::


# Foraging

::: {.panel-tabset}

## Distribution 

```{r}
hist1(data$Foraging)
table(data$Foraging==0)
descdist(data$Foraging, boot = 1000, print = FALSE)
```

The data are clearly not following a normal distribution. It seems that a beta distribution could suit but foraging is not a continuous variable ranging from 0 to 1. Foraging has thus been analysed as a binary variable: 1 if fish spent time foraging during this trial, 0 otherwise.

There were 168 trials (out of 252) during which no foraging happened.

## Plot 

The plots are in percentage of trials during which the fish foraged over the total number of trials.

```{r}
#| warning: false
plot_var_binary(data, "Foraging.bin")
plot_var_binary_fish(data, "Foraging.bin")
```

## Generalised linear model 

```{r}
#| column: page
mForaging <- glmmTMB(Foraging.bin ~ Tank + Time + Filter + Order + (1|Fish), data = data, family =binomial)

glmmTMB:::Anova.glmmTMB(mForaging)

contr_tank.size_bin(mForaging)
contr_tank.enrich_bin(mForaging)
# example of differences between Jar and Large
## Contrast estimate
1/0.09
## Upper CI
1/0.015
## Lower CI
1/0.493
```

Effect of Tank. 

As a reminder, an odds.ratio < 1 between tank X vs tank Y means the probability of foraging during a trial in tank X is smaller that the probability of foraging during a trial in tank Y. An odd ratio for instance of 0.2 means that a fish is (1 / 0.2) = five times more likely to forage during a trial in tank Y than in Tank X. If we imagine 5 trials, fish will forage on average during 1 trial out of 5 in Tank X whereas fish will forage on average during 5 trials out of 5 in Tank Y. 

+ Tank. Fish were more likely to forage during a trial in Large compared to Jar, Small and Medium. Again in accordance with the PCA and the results of Swimming and Resting. Fish in Large were on average 11 times [CI 2, 67] more likely to forage during a trial than fish in Jar, xx times [CI xx, xx] than fish in Small, and xx times [CI xx, xx] than fish in Medium. In addition, fish were xx times [CI xx, xx] more likely to forage during a trial in Large than Barren. In accordance with the PCA and the results of Swimming. This time, there was also a significant difference between small and medium.

## Between-fish diff

```{r}
rpt.foraging <- rpt(Foraging.bin ~ Tank + Time + Filter + Order + (1|Fish), grname = "Fish", data = data, datatype = "Binary", nboot = 0, npermut = 0)
# can calculate repeatability but not the CI. I tried using bootstrap and permutation.

# Likelihood ratio test of random effect
mForaging0 <- glmmTMB(Foraging.bin ~ Tank + Time + Filter + Order, data = data, family =binomial)
anova(mForaging0, mForaging)
```

Can calculate the repeatability but not its CI. Cannot calculate repeatability for Stereotypic swimming anyway. Hard to estimate repeatability on binary data. I don't know how to do it without the package rptR. I think we don't have enough data power anyway to calculate repeatability with a binary variable.

Likelihood ratio test indicates a statistically significant random effect. Meaning that fish behaved differently among each others in terms of foraging. 

## Diagnosis 

```{r}
mForaging_res <- simulateResiduals(mForaging)
plot(mForaging_res)
plotResiduals(mForaging_res, form = data$Tank)
plotResiduals(mForaging_res, form = data$Time)
plotResiduals(mForaging_res, form = data$Filter)
```

The model seems good.

:::

# Stereotypic swimming

::: {.panel-tabset}

## Distribution 

```{r}
hist1(data$Stereotypic.swimming)
table(data$Stereotypic.swimming==0)
descdist(data$Stereotypic.swimming, boot = 1000, print = FALSE)
```

Idem Foraging. Analysis as binary variable if the fish performs stereotypic swimming during a trial or not. 

There were 172 trials (out of 252) during which no stereotypic swimming happened.

## Plot 

```{r}
#| warning: false
(plotSS1 <- plot_var_binary(data, "SS.bin"))
(plotSS2 <- plot_var_binary_fish(data, "SS.bin"))
```

Huge variability between fish as always. Some fish never performed stereotypic swimming while some in all tanks. 

## Generalised linear model 

```{r}
#| column: page
mSS <- glmmTMB(SS.bin ~ Tank + Time + Filter + Order + (1|Fish), data = data, family =binomial)

glmmTMB:::Anova.glmmTMB(mSS)

contr_tank.size_bin(mSS)
contr_tank.enrich_bin(mSS)
```

Effect of Tank. 

+ Tank. Jar were less likely to perform stereotypic swimming during a trial than Small/Medium/Large but only significant between Jar and Small. Fish in Small were xx times [CI xx, xx] more likely to perform stereotypic swimming during a trial than fish in Jar. Almost significant difference between Large and Barren: Fish in Barren were 3.4 times [CI xx, xx] more likely to perform stereotypic swimming during a trial than fish in Large.

## Between-fish diff

```{r}
# Attempt to calculate repeatability
rpt.SS <- rpt(SS.bin ~ Tank + Time + Filter + Order + (1|Fish), grname = "Fish", data = data, datatype = "Binary", nboot = 0, npermut = 0)
# error

# Likelihood ratio test of random effect
mSS0 <- glmmTMB(SS.bin ~ Tank + Time + Filter, data = data, family =binomial)
anova(mSS0, mSS)
```

Error when calculating the repeatability with rptR. 

Likelihood ratio test indicates a statistically significant random effect. Meaning that fish behaved differently among each others in terms of stereotypic swimming. 

## Diagnosis

```{r}
mSS_res <- simulateResiduals(mSS)
plot(mSS_res)
plotResiduals(mSS_res, form = data$Tank)
plotResiduals(mSS_res, form = data$Time)
plotResiduals(mSS_res, form = data$Filter)
# Problem with Filter, I retried the model without Filter
mSS2 <- glmmTMB(SS.bin ~ Tank + Time + (1|Fish), data = data, family =binomial)
plot(simulateResiduals(mSS2))
```

It is expected to have a problem with filter as there are more than twice more trials without filter than with filter. No problem when removing Filter from the model. The model seems good.

## Stereotypic types

Just a plot with the type of stereotypic swimming depending on the tank. To stick with the analysis, it is a count (count = 1 if the fish performs a certain type of stereotypic swimming during a trial). 

```{r}
#| column: page
#| echo: false
data_SST <- tf(master[master$Stereotypic.swimming !=0 & !is.na(master$Stereotypic.swimming),], 
            c("Tank","Stereotypic.swimming.type","Fish","Filter","Time"), 
            c("Stereotypic.swimming","Swimming"), # I have to specify two columns otherwise my function tf is not working
            function(x) colSums(x, na.rm = TRUE))

ggplot(data_SST, aes(fill = Stereotypic.swimming.type, x = Tank))+
    geom_bar(stat= "count",color=grey(0.4))+
    theme(legend.position="bottom")+
    ylab("Count of trials")

# Grouping type of stereotypic swimming together
data_SST$Stereotypic.swimming.type2 <- factor(data_SST$Stereotypic.swimming.type, 
                                              labels = c("Circles", "Pacing", "Pacing", "Pacing","Pacing", "Zig Zag"))

ggplot(data_SST, aes(fill = Stereotypic.swimming.type2, x = Tank))+
    geom_bar(stat= "count",color=grey(0.4))+
    theme(legend.position="bottom")+
    ylab("Count of trials")
```

It seems that there are not differences between tanks. Just circles seem specific to larger tanks (Medium/Large/Barren).

## Stereotypic Swimming repetitions

```{r}
#| column: page
#| echo: false
data_SST2 <- tf(master[master$Stereotypic.swimming !=0 & !is.na(master$Stereotypic.swimming),], 
            c("Tank","Stereotypic.swimming.type","Fish","Filter","Time"), 
            c("Stereotypic.swimming.number.of.repititions","Swimming"), # I have to specify two columns otherwise my function tf is not working
            function(x) colSums(x, na.rm = TRUE))

hist(data_SST2$Stereotypic.swimming.number.of.repititions)

data_SST2 %>%
  mutate(Nb.rep = Stereotypic.swimming.number.of.repititions) %>% 
  dplyr::select(Tank, Fish, Time, Stereotypic.swimming.type, Nb.rep) %>% 
  arrange(Fish, Tank, Time)
```

I don't think there are enough data to really say something about the number of repetitions. The number of repetitions seems pretty unique to a fish in a certain tank. I don't see an obvious pattern depending on the tank. The interesting thing is that the stereotypic swimming seemed consistent in a tank for a fish (most or all trials of that day with stereotypic swimming). It is a pitty we couldn't calculate the repeatability of the binomial glm of Stereotypic Swimming because it would have been interesting. 

:::

# Interaction with Surface

::: {.panel-tabset}

## Distribution 

```{r}
hist1(data$Interation.with.surface)
table(data$Interation.with.surface==0)
```

Idem analysis binary generalised linear models.

## Plot 

```{r}
#| warning: false
(plotInteracting1 <- plot_var_binary(data, "Interacting.bin"))
(plotInteracting2 <- plot_var_binary_fish(data, "Interacting.bin"))
```

## Generalised linear model 

```{r}
#| column: page
mInteracting <- glmmTMB(Interacting.bin ~ Tank + Time + Filter + Order + (1|Fish), data = data, family =binomial)

glmmTMB:::Anova.glmmTMB(mInteracting)

contr_tank.size_bin(mInteracting)
contr_tank.enrich_bin(mInteracting)
```

Effect of Tank.

+ Tank. Fish were more likely to interact with the surface during a trial in Jar/Medium compared to Large (not significant for Small). Fish in Jar were x times [CI xx, xx] more likely to interact with the surface during a trial than fish in Large, and Fish in Medium were x times [CI xx, xx] then fish in Large. In addition, fish were xx times [CI xx, xx] more likely to interact with a surface during a trial in Barren than in Large. 

## Between-fish diff

```{r}
mInteracting0 <- glmmTMB(Interacting.bin ~ Tank + Time + Filter + Order, data = data, family =binomial)
anova(mInteracting0, mInteracting)
```

Likelihood ratio test indicates a statistically significant random effect. Meaning that fish behaved differently among each others in terms of interaction with surface. 

## Diagnosis

```{r}
mInteracting_res <- simulateResiduals(mInteracting)
plot(mInteracting_res)
plotResiduals(mInteracting_res, form = data$Tank)
```

The model seems good.

## Interacting types

Just a plot with the type of interaction depending on the tank. To stick with the analysis, it is a count (count = 1 if the fish performs a certain type of interaction with surface during a trial). 

The second plot groups similar interactions with surface together.

```{r}
#| column: page
#| echo: false
data_IT <- tf(master[master$Interation.with.surface !=0 & !is.na(master$Interation.with.surface),], 
            c("Tank","Interaction.with.surface.type","Fish","Filter","Time"), 
            c("Interation.with.surface","Swimming"), # I have to specify two columns otherwise my function tf is not working
            function(x) colSums(x, na.rm = TRUE))

ggplot(data_IT, aes(fill = Interaction.with.surface.type, x = Tank))+
    geom_bar(stat= "count",color=grey(0.4))+
    theme(legend.position="bottom")+
    ylab("Count of trials")

# grouping types of interaction together
data_IT$Interaction.with.surface.type2 <- factor(data_IT$Interaction.with.surface.type, 
                                              labels = c(rep("Aggressive",8), rep("Bite",2), "Fold", rep("Head into",4),rep("Head shake/swim",5), "Bite",rep("Swim",2)))

ggplot(data_IT, aes(fill = Interaction.with.surface.type2, x = Tank))+
    geom_bar(stat= "count",color=grey(0.4))+
    theme(legend.position="bottom")+
    ylab("Count of trials")
```

:::

# Nest.building

::: {.panel-tabset}

## Distribution 

```{r}
hist1(data$Nest.building)
table(data$Nest.building==0)
```

There are 1 trial out of 5 where there was nest building. Analysis with binary generalised model.

## Plot 

```{r}
#| warning: false
(plotNest1 <- plot_var_binary(data, "Nest.bin"))
(plotNest2<- plot_var_binary_fish(data, "Nest.bin"))
```

## Generalised linear model 

```{r}
#| column: page
mNest <- glmmTMB(Nest.bin ~ Tank + Time + Filter + Order + (1|Fish), data = data, family =binomial)

glmmTMB:::Anova.glmmTMB(mNest)

calc_contrasts_bin(mNest)
emmeans(mNest, ~ Filter, type = "response")
```

Effect of Filter only.
+ Tank. Almost significant.
+ Filter. Fish were more likely to build a nest without a filter in their tank.

## Between-fish diff

```{r}
mNest0 <- glmmTMB(Nest.bin ~ Tank + Time + Filter + Order, data = data, family =binomial)
anova(mNest0, mNest)
```

Likelihood ratio test indicates a statistically significant random effect. Meaning that fish behaved differently among each others in terms of nest building. 

## Diagnosis

```{r}
mNest_res <- simulateResiduals(mNest)
plot(mNest_res)
plotResiduals(mNest_res, form = data$Tank)
plotResiduals(mNest_res, form = data$Time)
plotResiduals(mNest_res, form = data$Filter)
```

The model seems good.

# Hovering

To analyse the amount of time the fish was hovering on the 204 trials with hovering time > 0, we tried different models described in "Analysis_Hovering". The best model was a linear mixed model on the amount of time spent hovering with weighted least squares (WLS). We had to use WLS because the error variance was increasing with the fitted values-higher error variance associated with high hovering time values. It was not possible to use WLS with the lmer package so we had to use the nlme package. The nlme package does not handle confounding levels of two factors - the fact that Tank = Barren is the same as Order = 5. So we had to remove Order from the analysis. The addition of Order was not changing the analysis outcomes in the other models. 

::: {.panel-tabset}

## Distribution 

```{r}
hist1(data$Hovering)
table(data$Hovering==0)
descdist(data$Hovering, boot = 1000, print = FALSE)
quantile(data$Hovering)
hist1(data$Hovering[data$Hovering!=0]) # histogram without the zeros
```

Even if there are less zeros than Foraging or Stereotypic Swimming, it is not a normal distribution. There were only 48 trials (out of 252) during which no hovering happened. So 204 trials where we could study the amount of time hovering in 204 trials. 

We fitted a glm distribution like Foraging and Stereotypic Swimming on the probability to hover (dependent variable is 1 if hovering occurred during the trial, zero otherwise).

## Plot 

```{r}
#| warning: false
# Effect of tank of the probability to hover during a trial
plot_var_binary(data, "Hovering.bin")
plot_var_binary_fish(data, "Hovering.bin")

# Among the trials where the fish hovers, effect of tank on the amount of time hovering
data %>% filter(Hovering >0) %>% 
  plot_var(., "Hovering")
data %>% filter(Hovering >0) %>% 
  plot_var_fish(., "Hovering")
```

Both the probability to hover and the time spent hovering seems to be higher in Jar compared to the others. All fish besides Goblin hovered more in Jar compared to Small.

## Generalised linear model 

Y variable = probability to hover during a trial

```{r}
#| column: page
mHovering <- glmmTMB(Hovering.bin ~ Tank + Time + Filter + Order + (1|Fish), data = data, family =binomial)

glmmTMB:::Anova.glmmTMB(mHovering)

emmeans(mHovering, ~ Time, type = "response")
emmeans(mHovering, ~ Filter, type = "response")
```

No effect of Tank on the probability to hover during a trial, only an effect of Time and Filter

+ Time. Fish were more likely to hover in the morning.
+ Filter. Fish were less likely to hover with a filter in their tank.

```{r}
mHovering_res <- simulateResiduals(mHovering)
plot(mHovering_res)
```

The model seems good. 

## Linear model

Y variable = amount of time spent hovering during a trial. Only trials with hovering time > 0

```{r}
mHovering2 <- lme(Hovering ~ Tank + Time + Filter, random = ~1 | Fish, data = data[data$Hovering>0, ], weights = varPower(), control = lmeControl(maxIter = 20000))
anova(mHovering2)
contr_tank.size(mHovering2)
contr_tank.enrich(mHovering2)
emmeans(mHovering2, ~ Time)
emmeans(mHovering2, ~ Filter)
```

Effect of Tank, Time and Filter.

+ Tank. Jar hovered significantly more time than Small, Medium and Large. Fish in Jar hovered 53 sec [CI 20, 85] more than Small (over a 600 sec trial), 53 sec [CI 20, 86] more than Medium, and 61 sec [CI 28, 93] more than Large. Fish in medium hovered more time than fish in Large. In addition, fish in Barren hovered significantly more time (22 sec [CI 12, 35]) than fish in Large.
+ Time. Fish hovered more time in the middle of the day.
+ Filter. Fish hovered less time with a filter in their tank.

```{r}
plot(mHovering2, resid(., type = "pearson") ~ fitted(.), aspect = 1, pch = 21, abline = 0)
plot(mHovering2, abs(resid(., type = "pearson")) ~ fitted(.), aspect = 1, pch = 21, abline = 0)
qqnorm(resid(mHovering2, type = "pearson"))
qqline(resid(mHovering2, type = "pearson"))
ranefPlot <- ranef(mHovering2, level = 1)
qqnorm(ranefPlot$`(Intercept)`)
qqline(ranefPlot$`(Intercept)`)
```

The model seems good besides the normality. But linear models are robust to the normality assumption (= we can still draw true inference).

## Hovering place

Just a plot with where the fish hovers at the different places depending on the tank.

```{r}
#| column: page
#| echo: false
#| message: false
data_HP %>% 
  filter(!is.na(Hovering.place)) %>% 
  group_by(Tank, Hovering.place) %>% 
  summarise(Hovering = mean(Hovering)) %>% 
  ggplot(., aes(y = Hovering, x = Tank,fill = Hovering.place))+
    geom_bar(position = position_fill(reverse = TRUE), stat = "identity", color = "black")+
  scale_fill_manual(values=palette_HP)+
  guides(fill = guide_legend(reverse = TRUE))

data_HP %>% 
  filter(!is.na(Hovering.place)) %>% 
  group_by(Fish, Hovering.place) %>% 
  summarise(Hovering = mean(Hovering)) %>% 
  ggplot(., aes(y = Hovering, x = Fish,fill = Hovering.place))+
    geom_bar(position = position_fill(reverse = TRUE), stat = "identity", color = "black")+
  scale_fill_manual(values=palette_HP)+
  guides(fill = guide_legend(reverse = TRUE))
```

Fish used different places for hovering in all tanks.

:::

:::

# Up and down 

::: {.panel-tabset}

Research question: did fish use differently use the upper part or lower part of their tank depending of the tank type?

## Plot
```{r}
ggplot(dataUpDown, aes(y = adj.up, x = Tank)) +
    geom_jitter(alpha = .5, width = 0.05, height=0, color = grey(0.25)) +
    stat_summary(fun.y = mean,geom = "point",colour = "red", size = 4)+
    stat_summary(fun.min = function(x) mean(x) - sd(x), 
                 fun.max = function(x) mean(x) + sd(x), 
                 geom = "errorbar",colour = "red", width = 0.15)+
  ylab("Time up (sec)")
```

It seems that the tank has an influence on whether fish prefer being on the upper half or lower half of the tank.

## Model

We can do different models:
+ Look at the percentage of time spent up of the trial. Fit a beta distribution. Problem: All percentages need to be 0 < perc < 100%. There are trials with 100% so we can't use beta distribution. 

+ Look at the count of seconds (how many seconds the fish spend up of the tank). Fit a binomial distribution. Problem: The diagnostics plots were very bad when fitting mUp <- glmmTMB(cbind(up, down) ~ Tank + Time + Filter + (1|Fish), data = dataUpDown, family = binomial)

+ Look simply at the actual time spent up. Fit a normal distribution. The response variable can be adjusted by the fact that the trial were not perfectly 600 sec. $adj.up= \frac{\text{time spent up} * 600}{\text{trial total duration}}$. It is what has been done.

```{r}
mUp <- lmer(adj.up ~ Tank + Time + Filter + Order + (1|Fish), data = dataUpDown)
anova(mUp, ddf = "Kenward-Roger", type = 2)
contr_tank.size(mUp)
contr_tank.enrich(mUp)
emmeans(mUp, ~ Time) # estimated means of time spent swimming depending on the time of the day 
emmeans(mUp, ~ Filter) # estimated means of time spent swimming depending on the time of the day
```

Effect of Tank, Time and Filter. 

+ Tank. Statistically significant difference between Small and Medium/Jar: Fish in Small tank spent on average 103 more seconds [95% CI 26, 181] on the upper half of the tank compared to fish in Medium tanks. At the opposite, Fish in Small tank spent on average 71 more seconds [95% CI 26, 181] on the lower half of the tank compared to fish in Jar tanks. 
+ Time. Fish spent more time on the upper half of the tank in the morning.
+ Filter. Fish spent more time on the upper half of the tank with a filter in their tank.

## Diagnosis

```{r}
lm_diagnosis(mUp)
```

The model seems good.

:::

# R version 

```{r}
sessionInfo()
```

:::
