---
title: "Order of tanks"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
# Run the scripts that load the data and that contains the functions to plot and do the linear models
source("Script/Load-data.R")
source("Script/Functions/Plots.R")
source("Script/Functions/Linear-model.R")
data$Order.fac <- factor(data$Order)
```

Order in which the fish experienced the tanks. Fish underwent different order of treatments: Some fish started with Large, then Small, etc; Other fish with Medium, then Jar, etc. Barren was the last tank = 5th tank for all fish. Is there an effect of the order of treatments? For instance, were fish more swimming on average during the 1st or 2nd tank they went in?  

A problem because all fish last experienced Barren, we cannot disentangle the effect of Barren from the effect of order. Maybe all the changes in behaviour in Barren were only because it was the last tank the fish went into. At least, we are comparing Large and Barren -Large is not always the 4th tank- so we are not always comparing the same tank orders when comparing Large and Barren. 

```{r}
#| warning: false
data %>% filter(Tank != "Barren") %>% 
  plot1("Order", main = .) +
  ggtitle("Tank order")
```

# Swimming

Treat tank order as a continuous or categorical variable? 

```{r}
plot(data$Swimming ~ data$Order.fac)
ggplot(data, aes(y = Swimming, x = Order)) + geom_point() + geom_smooth()
```

Considering the plot of the relationship between Swimming and Tank order (no linear relationship), tank order is better as a categorical variable.

After, comparison of models with or without tank order, on the full or subset dataset without Barren

```{r}
# With tank order on data without Barren
mOrder.Swimming.subset <- lmer(Swimming ~ Tank + Time + Filter + Order.fac + (1|Fish), data = filter(data, Tank !="Barren"))
anova(mOrder.Swimming.subset , ddf = "Kenward-Roger", type = 2) 
emmeans(mOrder.Swimming.subset, ~ Tank) %>% 
  contrast(.,
           list(jar_vs_small     = c(1,-1, 0, 0), 
                jar_vs_medium    = c(1, 0,-1, 0), 
                jar_vs_large     = c(1, 0, 0,-1),
                small_vs_medium  = c(0, 1,-1, 0),
                small_vs_large   = c(0, 1, 0,-1), 
                medium_vs_large  = c(0, 0, 1,-1)),
           adjust = "holm")
# Without tank order on data without Barren
mSwimming.subset <-lmer(Swimming ~ Tank + Time + Filter + (1|Fish), data = filter(data, Tank !="Barren")) 
anova(mSwimming.subset , ddf = "Kenward-Roger", type = 2)
AIC(mOrder.Swimming.subset , mSwimming.subset)

# With tank order on all data
mOrder.Swimming <-lmer(Swimming ~ Tank + Time + Filter + Order.fac + (1|Fish), data = data) 
anova(mOrder.Swimming , ddf = "Kenward-Roger", type = 2)
## Contrasts among Jar, Small, Medium and Large over Orders 1 to 4
emmeans(mOrder.Swimming, ~ Tank , 
                       at = list(Tank = c("Jar", "Small", "Medium", "Large"), 
                                 Order.fac = c("1", "2", "3", "4"))) %>% 
  contrast(.,
           list(jar_vs_small     = c(1,-1, 0, 0), 
                jar_vs_medium    = c(1, 0,-1, 0), 
                jar_vs_large     = c(1, 0, 0,-1),
                small_vs_medium  = c(0, 1,-1, 0),
                small_vs_large   = c(0, 1, 0,-1), 
                medium_vs_large  = c(0, 0, 1,-1)),
           adjust = "holm")
## Contrasts between Large (over Orders 1 to 4) and Barren
emm1 <- emmeans(mOrder.Swimming, ~ Tank , 
                       at = list(Tank = c("Large"), 
                                 Order.fac = c("1", "2", "3", "4")))
emm2 <- emmeans(mOrder.Swimming, ~ Tank , 
                       at = list(Tank = c("Barren"), 
                                 Order.fac = c("5")))
contrast(rbind(emm1,emm2), list(large_vs_barren= c(1,-1)))
## Estimated means for Orders 1 to 4
emm1 <- emmeans(mOrder.Swimming, ~ Order.fac, 
                       at = list(Order.fac = c("1", "2", "3", "4")))

# Without tank order on all data
mSwimming <-lmer(Swimming ~ Tank + Time + Filter + (1|Fish), data = data) 

AIC(mOrder.Swimming,mSwimming)
```

Significant effect of tank order + The best model is the one with tank order (in terms of AIC) on the full dataset or dataset without Barren => Tank order needs to be included in the model... 

Good news:
+ I found a way to calculate contrasts among Jar, Small, Medium and Large on the model with all data!
+ I found a way to calculate the contrast between contrast Large (estimated across Orders 1 to 4) and Barren at Order 5.
+ Still an effect of tank when tank order is included

# Resting

```{r}
# with tank order
mOrder.Resting <- lmer(Resting ~ Tank + Time + Filter + Order.fac + (1|Fish), data = data)
anova(mOrder.Resting, ddf = "Kenward-Roger", type = 2) 

# without tank order
mResting <- lmer(Resting ~ Tank + Time + Filter + (1|Fish), data = data) 

AIC(mOrder.Resting, mResting)
```

Idem Swimming

# Foraging

```{r}
mOrder.Foraging <- glmmTMB(Foraging.bin ~ Tank + Time + Filter + Order.fac + (1|Fish), data = data, family =binomial, control = glmmTMBControl(rank_check = "adjust")) 
glmmTMB:::Anova.glmmTMB(mOrder.Foraging)

mForaging <- glmmTMB(Foraging.bin ~ Tank + Time + Filter + (1|Fish), data = data, family =binomial)

AIC(mOrder.Foraging,mForaging)
```

The model is better without Tank order + no effect of tank order. Still a significant effect of Tank with Tank order in the model.

# Nest building

```{r}
mOrder.Nesting <- glmmTMB(Nest.bin ~ Tank + Time + Filter + Order.fac + (1|Fish), data = data, family =binomial, control = glmmTMBControl(rank_check = "adjust")) 
glmmTMB:::Anova.glmmTMB(mOrder.Nesting)

mNesting <- glmmTMB(Nest.bin ~ Tank + Time + Filter + (1|Fish), data = data, family =binomial)

AIC(mOrder.Nesting,mNesting)
```

Idem Foraging, the model is better without Tank order + no effect of tank order. However, keeping Tank order removes the significant effect of Tank on Nest Building

# Stereotypic Swimming

```{r}
mOrder.SS <- glmmTMB(SS.bin ~ Tank + Time + Filter + Order.fac + (1|Fish), data = data, family =binomial, control = glmmTMBControl(rank_check = "adjust")) 
glmmTMB:::Anova.glmmTMB(mOrder.SS)

mSS <- glmmTMB(SS.bin ~ Tank + Time + Filter + (1|Fish), data = data, family =binomial)

AIC(mOrder.SS,mSS)
```

Idem Foraging

# Dragon

What happen to Dragon during its exception of the first tank it went into? Do we consider the second tank it went into as tank order = 1 or tank order = 2? 

# Conclusion

I think we need to include Tank order in the model... A bit of a pain. It took me a long time to figure out how to calculate estimated means and contrasts on the new models that are rank-deficient factorial ANOVA, but I did it! 